OpenMPIR

### 2017-10-08 paper.tex
fix overlapping text in header
Added uts footnote

@@

To evaluate the use of Tapir we compared performance on the Barcelona OpenMP Task 
Suite to several existing OpenMP implementations. The suite is a set of tests
intended to test performance of OpenMP implementations using both irregular and
regular tasking. The following list provides abridged descriptions of each
benchmark taken verbatim from the test suite distribution \cite{barcelona}.

@@

@@

The surprising fact that program behavior wasn't changed significantly is worth
discussing. We surmise that it is likely due to the semantics of variables
borrowed from the Cilk codegen and their relation to the OpenMP specification
of semantics. In particular, \emph{OpenMP specifies a relaxed memory model
for shared variables}, in which writes by one thread need only be visible at
to other threads at explicit or implicit flushes. 
This results in the Cilk implementation of only writing
to a shared variable at the end of a parallel section being a valid
implementation of OpenMP's specification. In contrast, existing OpenMP
implementations sometimes write shared variables to memory on every write access. This is a
potential for performance discrepancy because of differing 
implementation behavior.
%Stephen:This topic is not discuessed later, but was touched on before -- moved?
%and we will return to it when discussing
%performance on the Barcelona OpenMP task suite. 
There is of course the fact
that technically, by copying the value of variables back to the surrounding
context even for \texttt{private} variables, one isn't following the
specification. We surmise that in practice this hasn't had an effect because
\texttt{private} variables are often used for inputs to calculations that are 
not accessed after the tasks complete, 
or for performance rather than their semantic properties.
Also, \texttt{firstprivate} is the default data-sharing attribute for tasks.

@@

One of the important questions in this work is whether the Tapir
instructions can be extended to cover the full OpenMP semantics.
%As we have shown, even the full semantics of OpenMP tasks weren't covered.
Runtime
calls aside, implementing the full semantics of OpenMP pragmas would be a
significant undertaking.

>>>>

\renewcommand{\shortauthors}{G. Stelle et al.}

@@

To evaluate the use of Tapir we compared performance on the Barcelona OpenMP Task 
Suite to several existing OpenMP implementations. The suite is a set of tests
intended to test performance of OpenMP implementations using both irregular and
regular tasking.\footnote{Since original publication, an unbalanced tree search
benchmark has been added to the benchmark suite. Unfortunately, due to time
constraints, we weren't able get results for this additional benchmark.} The
following list provides abridged descriptions of each benchmark taken verbatim
from the test suite distribution \cite{barcelona}.

@@

Another way we can discriminate the cause of performance discrepency is by disabling 
Tapir optimizations. This allows us to directly measure the impact that Tapir
optimizations are having on runtime. For \texttt{fibonacci}, for example,
disabling optimizations results in a roughly ~2x performance loss. This is likely

@@

The surprising fact that program behavior wasn't changed significantly is worth
discussing. We surmise that it is likely due to the semantics of variables
borrowed from the Cilk codegen and their relation to the OpenMP specification
of semantics. In particular, \emph{OpenMP specifies a relaxed memory model
for shared variables}, in which writes by one thread need only be visible at
to other threads at explicit or implicit flushes. 
This results in the Cilk implementation of only writing
to a shared variable at the end of a parallel section being a valid
implementation of OpenMP's specification. In contrast, existing OpenMP
implementations sometimes write shared variables to memory on every write
access. This is a potential for performance discrepancy because of differing
implementation behavior.
%Stephen:This topic is not discuessed later, but was touched on before --
%moved?  and we will return to it when discussing performance on the Barcelona
%OpenMP task suite. 
There is of course the fact that technically, by copying the value of variables
back to the surrounding context even for \texttt{private} variables, one isn't
following the specification. We surmise that in practice this hasn't had an
effect because \texttt{private} variables are often used for inputs to
calculations that are not accessed after the tasks complete, or for performance
rather than their semantic properties.  Also, \texttt{firstprivate} is the
default data-sharing attribute for tasks.

@@

One of the important questions in this work is whether the Tapir instructions
can be used to implement the full OpenMP semantics.
%As we have shown, even the full semantics of OpenMP tasks weren't covered.
Runtime calls aside, implementing the full semantics of OpenMP pragmas would be
a significant undertaking.
